Non-Modal Predictions of Turbulent Properties Applied to the Hasegawa-Wakatani Model

### 2014-12-24 Referee_Responses/referee_response.tex

\documentclass[12pt]{article}
\usepackage{graphicx}
\setlength{\parindent}{0in}

% General Latex  --------------------------------------------------
\def\beq{\begin{equation}}
\def\eeq{\end{equation}}
\def\beqar{\begin{eqnarray}}
\def\eeqar{\end{eqnarray}}
\def\nn{\nonumber}
\def\ol{\overline}
\def\para{\parallel}

% Operators  ------------------------------------------------------
\newcommand{\diff}[2]{\frac{d#1}{d#2}}
\newcommand{\diffs}[2]{\frac{d^2#1}{d#2^2}}
\newcommand{\pdiff}[2]{\frac{\partial#1}{\partial#2}}
\newcommand{\pdiffs}[2]{\frac{\partial^2#1}{\partial#2^2}}
\newcommand{\pdiffxy}[3]{\frac{\partial^2#1}{\partial#2 \partial#3}}
\newcommand{\pdt}{\partial_t}
\newcommand{\pdr}{\partial_r}
\newcommand{\pdth}{\partial_\theta}
\newcommand{\pdrr}{\partial^2_r}

\newcommand{\enum}[2]{{#1}\times10^{#2}} % 4.2x10^{3} = \enum{4.2}{3}

\newcommand{\vect}[1]{{\bf #1}}
%\newcommand{\vect}{\overrightarrow}
%\newcommand{\vect}{\vec}
\def\div{\nabla\cdot}
\def\grad{\nabla}
\def\curl{\nabla\times}
\newcommand{\gradpar}{\grad_\parallel}
\newcommand{\gradperp}{\grad_\perp}
\newcommand{\gradr}{\grad_r}
\newcommand{\defeq}{\ensuremath{\stackrel{\text{\tiny def}}{=}}}

\newcommand{\savg}[1]{\left<{#1}\right>}
\newcommand{\vavg}[1]{\left<{#1}\right>_V}
\newcommand{\thavg}[1]{\left<{#1}\right>_\theta}

% Variable names  -------------------------------------------------
\newcommand{\vpar} {v_\parallel}
\newcommand{\Apar} {A_\parallel}
\newcommand{\jpar} {j_\parallel}
\newcommand{\kpar} {k_\parallel}
\newcommand{\kperp} {k_\perp }
\newcommand{\vperp} {v_\perp }
\newcommand{\kthe}{k_\theta}

\newcommand{\Evec}{\ensuremath{\boldsymbol{{\rm E}}}}
\newcommand{\Bvec}{\ensuremath{\boldsymbol{{\rm B}}}}
\newcommand{\Jvec}{\ensuremath{\boldsymbol{{\rm J}}}}
\newcommand{\Fvec}{\ensuremath{\boldsymbol{{\rm F}}}}
\newcommand{\fvec}{\ensuremath{\boldsymbol{{\rm f}}}}
\newcommand{\vE}{\ensuremath{\boldsymbol{{\rm v}_{E}}}}
\newcommand{\bo}{\ensuremath{\boldsymbol{{\rm b}_0}}}
\newcommand{\bvec}{\ensuremath{\boldsymbol{{\rm b}}}}
\newcommand{\xvec}{\ensuremath{\boldsymbol{{\rm x}}}}
\newcommand{\yvec}{\ensuremath{\boldsymbol{{\rm y}}}}
\newcommand{\zvec}{\ensuremath{\boldsymbol{{\rm z}}}}
\newcommand{\vvec}{\ensuremath{\boldsymbol{{\rm v}}}}
\newcommand{\jvec}{\ensuremath{\boldsymbol{{\rm j}}}}

\newcommand{\bxgp}{\bvec\times\gradperp}

\newcommand{\vve}{\ensuremath{\boldsymbol{{\rm v}}_{e}}}
\newcommand{\vvi}{\ensuremath{\boldsymbol{{\rm v}}_{i}}}
\newcommand{\vpe}{v_{\parallel e}}
\newcommand{\vpi}{v_{\parallel i}}
\newcommand{\vvE}{\ensuremath{\boldsymbol{{\rm v}}_{E}}}
\newcommand{\vvD}{\ensuremath{\boldsymbol{{\rm v}}_{D}}}

\newcommand{\nuei}{\nu_{ei}}
\newcommand{\nuii}{\nu_{ii}}
\newcommand{\nue}{\nu_{e}}
\newcommand{\nuen}{\nu_{en}}
\newcommand{\nuin}{\nu_{in}}
\newcommand{\kpe}{\kappa_{\parallel e}}

\newcommand{\rs}{\rho_{s}}
\newcommand{\ri}{\rho_{i}}
\newcommand{\wci}{\Omega_{i}}
\newcommand{\wcix}{\Omega_{ix}}
\newcommand{\wce}{\Omega_{e}}
\newcommand{\tomega}{\tilde\omega}
\newcommand{\Isat}{I_{\rm sat}}
\newcommand{\fmie}{\frac{m_i}{m_e}}
\newcommand{\fmei}{\frac{m_e}{m_i}}

% Often used dimensions
\newcommand{\cm}{\rm cm}
\newcommand{\mm}{\rm mm}
\newcommand{\cmn}{{\rm cm}^{-3}}
\newcommand{\mn}{{\rm m}^{-3}}
\newcommand{\eV}{\rm eV}
\newcommand{\G}{\rm G}
\newcommand{\T}{\rm T}

\begin{document}

{\bf Response to Referees \\
 \\} 
Brett Friedman and Troy Carter
%Updated: \today \\

\hrulefill

We would like to thank the referee for his/her thorough reading of our paper and clear and insightful comments. We have responded to each comment below and made requisite changes to the text.
Most of the changes are relatively minor and involve corrections pointed out by the referee, changes in wording, and further clarification on certain points.

The referee's comment 14 is a rather significant one, suggesting the possible removal of Section V of the paper. 
We agree with the points but don't want to go so far as to remove the entire section because even though the result is weak (unsuccessful), 
we believe that it is still valuable and provides some important insight. What we have done, then, is move it to an appendix. 
We have also title, abstract, and introduction to de-emphasize that part of the paper. \\ \\

\emph{1) page 2, second column: $N_k$ is termed 'the nonlinear matrix'. Since matrices are generally associated to linear operator, it would be more appropriate to call it nonlinear operator. } \\

This is a good point. We have made this change. \\

\emph{2) page 3, Eq. 13. This is defined as an energy. However, it is a normalized (adimensional) quantity, and as such it looks more like an energy amplification factor.}\\

This is correct. We did not do a good job with our explanation. Really, we always set the initial amplitude to 1 so that the energy and the energy amplification are synonymous.
We have added this explanation in the text and changed Eq. 13. \\

\emph{3) page 3, 2nd column. The argument about non-normality of an operator implicitly assumes the choice of a norm (in this case $L_2$ norm, I presume). One can always choose a norm that makes a non-normal matrix normal, so the particular choice (and the physical reason behind it) should be spelled out. }\\

We did try to spell out what norm we were using around Eqs. 6-8, but I guess we didn't make it clear enough. So we have changed and added some text to try to make it clearer. We are using the $L_2$ norm. However, we have redefined our state vector (and the linear and nonlinear operators for consistency) so that the $L_2$ norm of the state vector gives the energy. This is a common thing that is done in many of our references.\\

\emph{4) page 4: the derivation of Eq. 17 is confusing. It is probably just a wrong equation reference ("we multiply eq. 15 and 16..."). I guess the procedure is simply to substitute eqs. 15 and 16 in 1 and 2, use the basis orthogonality and sum the two equations together. }\\

Yes, we did just reference the wrong equations. Sorry.\\

\emph{5) From section 3 on, several growth rate 'gamma' are introduced (with different subscripts). I suggest to list all of them in a table, with an explanation of their meaning. Otherwise it is too easy to get lost. }\\

Good idea. We've added a table now.\\

\emph{6) I am a bit confused by the use of eq. 18. The authors say that this is the way the energy evolution equation can be 'symbolically' (I suggest "formally") written. However, by inspection of eq. 17, one can explicitly write down an equation for the linear term $dE_l/dt$. Is not only the third term on the RHS of eq. 17 linear in the energy? In other words, it seems to me that the first and second terms on RHS cannot be written as linear combinations of the energy. This automatically gives the linear rate of energy injection $\gamma_{l,k}$. }\\

We attempt to make some clarifications in the text, namely by explicitly writing the equations for $dE_l/dt$ and $dE_{nl}/dt$. We do not mean that certain terms on the RHS
can be written as linear combinations of the energy. We mean that the first three terms come from the linear terms of the original HW equations. 
They can be used to evaluate the turbulent growth rate, which is somewhat analogous to the eigenmode growth rate. For instance, if $n_k, \phi_k$ are taken from one of the linear
eigenmodes and inserted into the first three terms of Eq. 17, and then one divides by (twice) the energy (Eq. 6) using the same $n_k, \phi_k$, one recovers the growth rate of that eigenmode.\\

\emph{7) page 4, last paragraph. "$\gamma_{t,k}<\gamma_{s,k}$". Here the authors are using $\gamma_t$ in lieu of $\gamma_l$, as explained in the previous paragraphs. However, this is legitimate only in quasi-steady state. Otherwise, $\gamma_t$ is a nonlinear quantity that changes in time, and can be larger than $\gamma_s$ even for systems with normal linear operators. }\\

We clarify in the text that $\gamma_t$ is defined in quasi-steady state. It is time-independent, as are all of the growth rates we define.\\

\emph{8) On page 5 the authors compare plots of $\gamma_s$ and $\gamma_t$, but they have not yet explained how these are calculated. I suggest to postpone any discussion about growth rate behavior, after explaining how they are calculated. }\\

The calculations of these growth rates are explained before the plots, and now they are hopefully explained better.\\

\emph{9) Page 5, first column: "moderate normality" should be "moderate non-normality" (i.e. normality is only for the Henrici measure equal to 0, while non-normality can be moderate, heavy, etc.) }\\

Good catch.\\

\emph{10) I have found quite confusing the explanation of how the turbulent growth rate $\gamma_t$ is calculated. On page 6, first column, it says: "may be approximated without large numerical simulations", which suggests that one needs to run small simulations. It turns out that, on the contrary, it can be evaluated analytically. However, there is a refrain in the paper referring to the complex ratio $n_k/\phi_k$. But I don't see where and why this needs to be calculated. }\\

We hope the edits have made this clear in the paper. $\gamma_t$ is calculated from the nonlinear simulation. Its calculation requires $n_k$ and $\phi_k$ as functions of time, although only the ratio $n_k/\phi_k$ is actually necessary. Our analytic method gives us a non-modal growth rate, which is an \emph{approximation} (good or bad) of the turbulent growth rate. So there is no need for small simulations, although we see how that was implied and we have thus removed the word ``large.''\\

\emph{11) Eq. 21 misses a factor 1/2 }\\

We had meant to define $u_k = 1/ \sqrt{2} (n_k, k \phi_k)^T$ so that the factor of $1/2$ wasn't necessary, but we missed the $\sqrt{2}$ factor in an earlier expression, so that was the incorrect one.\\

\emph{12) On Eq. 22. There is an unspelled assumption here on the distribution of the ensemble. Are all states equally distributed? Is such distribution physical meaningful? Also, I would emphasize the fact that by averaging over an ensemble of states, the information about the initial condition is lost (or averaged out). This is a crucial point because one of the main weaknesses of non-modal analysis with respect to the standard normal mode analysis, is the need to specify a given initial condition. }\\

Yes all states are equally distributed initially. Remember that each state at a particular $k$ is completely classified by the ratio $n_k/\phi_k$ as far as the growth rate is concerned.
So each member of the ensemble is initially one possible value of this complex ratio. But once the initial states have evolved under the linear operator for some time, not all of
the possible values of the ratio are populated. In fact, after a long enough time, this ratio is nearly equal to that for the faster growing eigenmode for all members of the ensemble.
The time average, then, produces something meaningful -- that is, an average over many ratios between a random ratio and the eigenmode ratio.

In physical terms, the system starts out random, then evolves toward the faster growing linear eigenmode structure until it reaches a nonlinear decorrelation time, at which point it becomes
random again. Remember, though, that we use an ensemble of initially random states in place of a long time series that involves randomization followed by linear evolution followed by
randomization, etc.

And yes, all information about the initial condition is lost, which is exactly our intention. Having to specify an initial condition is the main weakness of traditional non-modal analysis.
Our analysis uses ideas from non-modal analysis but attempts to average out the initial condition, making it more analogous to normal-mode analysis. So maybe when we say non-modal analysis,
we give the wrong impression, but we have no better term.\\

\emph{13) About Figure 6 and 7: these represent probably the most crucial figures of the paper. Yet, there is still no explanation on how $\gamma_t$ is evaluated from the full nonlinear simulations. }\\

Again, we've clarified it in the paper now.\\

\emph{14) I am very perplexed about the whole section V, to the extent that I would probably take it out. I appreciate the effort of using the discussed non-modal growth rate as indicator of subcritical turbulence onset. But the results of this section are very weak. First, the author have to introduce a yet different model from the ones previously described, in order to get something that can actually sustain subcritical turbulence. And more importantly, figure 11 suggest that their method does not work! So I don't understand why one would like to read this section. }\\

We agree that the section overall is weak due to the lack of success of the method in the subcritical regime, and the section isn't essential for the rest of the paper. But the issue
of predicting the subcritical onset is such an important one that it merits discussion. And although our particular method does not work, it seems plausible that some of the ideas
we use may be helpful in crafting a better method. And from doing the subcritical work, we have learned where the method breaks down. This information shouldn't be abandoned.
So we think turning Section V into an appendix may be the best solution to keep it from muddying the rest of the paper.\\

\emph{15) In figure 10a) I think it is more interesting to show the ratio of fluctuation over total energy, otherwise the numbers do not convey any information. }\\

Well the numbers are not that important, but this energy seems to be the natural normalized one: $|n|^2 + |\gradperp \phi|^2$. So we're not sure what you mean. \\

\emph{16) Finally, this manuscript contain too many typos!! Please run a spell-check before submitting (and, possibly, enumerate lines). }\\

We've now run spell-check and tried to catch any typos.

\bibliographystyle{unsrt}
\bibliography{refs}

%
%

\end{document}

### 2014-12-24 HW_non-modal.tex
Dec 23

\title{A Procedure to Predict the Subcritical Turbulent Onset Criterion Applied to a Modified Hasegawa-Wakatani Model}

@@

Non-modal analysis generally employs tools such as pseudospectra and maximum growth curves, which have been used to successfully explain
why turbulence is subcritical in various instances~\cite{trefethen1993,camargo1998}. 
However, they have not been used to predict onset criteria in subcritical systems (e.g. the transition Reynold's number in pipe flows).
This paper outlines and illustrates a non-modal method to make quantitative predictions of turbulent properties including turbulent onset criteria. 
Our approach is to define a non-modal growth rate that is analagous to an eigenmode growth rate, but which takes into account transient growth effects.

@@

In the analysis and numerical simulations, which we perform with the BOUT++ code~\cite{dudson2009}, we use periodic boundary conditions in
the $y$ and $z$ directions with zero-value boundaries in the $x$ direction. We use a domain size of $64 \rho_s \times 64 \rho_s \times 320 \rho_s$.
For the 2D version of the model, we replace $-\xi \gradpar^2$ with a single parameter $\alpha$ (definied as $\alpha = 4 \pi k_z^2 \rho_s^2 \omega_{ce} /\nu_{ei})$, 
which is commonly known as the adiabaticity parameter~\cite{camargo1995,camargo1998}. Additionally we use a domain size of $50 \rho_s \times 50 \rho_s$.
In numerical simulations, the turbulence can be suppressed due to the build-up of $k_y=0$ density and potential structures~\cite{biskamp1995}. To prevent this, we subtract out the $k_y=0$
density and potential components, which is the equivalent of adding source terms to Eqs.~\ref{n_eq} and~\ref{phi_eq}~\cite{friedman2012b}.

@@

\beqar
\label{fourier_eqn}
\frac{\partial}{\partial t} \left( \begin{array}{cc} n_k \\ \phi_k \end{array} \right) = \mathbf{A}_k \left( \begin{array}{cc} n_k \\ \phi_k \end{array} \right) + \mathbf{N}_k \left( \begin{array}{cc} n_k \\ \phi_k \end{array} \right), \\ \nonumber \\
\label{A_k}
\mathbf{A}_k = \left( \begin{array}{cc} -\xi k_z^2 - D k_\perp^4 & -i \kappa k_y + \xi k_z^2 \\  \xi \frac{k_z^2}{k_\perp^2} & - \xi \frac{k_z^2}{k_\perp^2} - D k_\perp^4\end{array} \right)
\eeqar
with $\mathbf{A}_k$ being the linear coupling matrix and $\mathbf{N}_k$ the nonlinear matrix, which we don't write explicitly. 
Now in assessing the degree of normality of the HW model, we are not interested in the normality of $\mathbf{A}_k$ itself. 
Rather, we are interested in the energetically-weighted version of this matrix~\cite{camargo1998,schmid2007,camporeale2010}. Since the energy in a given Fourier mode is

@@

\beqar
\label{fourier_en_eqn}
\frac{\partial}{\partial t} \left( \begin{array}{cc} n_k \\ k_\perp \phi_k \end{array} \right) = \mathbf{B}_k \left( \begin{array}{cc} n_k \\ k_\perp \phi_k \end{array} \right) + \mathbf{N}_k \left( \begin{array}{cc} n_k \\ k_\perp \phi_k \end{array} \right), \\ \nonumber \\
\label{B_k}
\mathbf{B}_k = \left( \begin{array}{cc} -\xi k_z^2 - D k_\perp^4 & -i \kappa \frac{k_y}{k_\perp} + \xi \frac{k_z^2}{k_\perp} \\  \xi \frac{k_z^2}{k_\perp} & - \xi \frac{k_z^2}{k_\perp^2} - D k_\perp^4\end{array} \right)
\eeqar
Since the square $L_2$-norm of the new state vector $\frac{1}{\sqrt{2}} \left( \begin{array}{cc} n_k \\ k_\perp \phi_k \end{array} \right)$ gives the energy contained in the mode $k$,
$\mathbf{B}_k$ is the energetically-weighted matrix of interest. In the adiabatic limit (as long as $k_\perp$ remains finite):

@@

\beq
\label{lin_HW}
\diff{u_k}{t} = \mathbf{B}_k u_k
\eeq
where $u_k = \left( n_k, k_\perp \phi_k \right)^T$.
The solution to this is $u_k(t) = e^{\mathbf{B}_k t} u_k(0)$ where the exponential of the matrix is defined in terms of
its Taylor expansion. The solution is dependent on the initial condition $u_k(0)$. The energy as a function of time is

\beq
\label{Et}
E_k(t) = \frac{||u_k(t)||^2}{||u_k(0)||^2} = \frac{||e^{\mathbf{B}_k t} u_k(0)||^2}{||u_k(0)||^2}
\eeq
Different initial conditions provide optimal energy growth at different times, but the upper envelope of all such energy evolution curves is given by
$E_{k,\rm{max}}(t) = ||e^{\mathbf{B}_k t}||^2$, which does not depend on an initial condition.
Moreover, $||e^{\mathbf{B}_k t}||^2 \ge e^{2 \gamma_{s,k} t}$, where $\gamma_{s,k}$ is the real part of
the fastest growing (or least damped) eigenmode of $\mathbf{B}_k$ (and of $\mathbf{A}_k$, since $\mathbf{B}_k$ is just a similarity transform of $\mathbf{A}_k$~\cite{camargo1998}). 
The equality holds when $\mathbf{B}_k$ is normal. 
Thus, non-normality in practice allows a system to undergo more energy growth than is dictated by the most unstable linear eigenmode. We show this in Fig.~\ref{max_en_growth} where we plot
$||e^{\mathbf{B}_k t}||^2$ for different values of $\alpha$ along with $e^{2 \gamma_{s,k} t}$, showing that transient non-modal growth is strong for small values of the scalar normality measure $1-H$, which scales monotonically
with the adiabaticity parameter $\alpha$.

\begin{figure}
\centerline{\includegraphics[width=0.45\textwidth]{max_en_growth}}
\caption{$G_{k,\alpha,\rm{max}}(t) = ||e^{\mathbf{B}_{k,\alpha} t}||^2$ for $k_x=0, k_y=1, \kappa=1$ for different values of the 2D adiabaticity parameter 
$\alpha$ compared to the energy growth $e^{2 \gamma_{s,\alpha,k} t}$ of the least stable eigenmode of  $\mathbf{B}_{k,\alpha}$. Also indicated for each matrix is its scalar normality measure $1-H$.
The time axis is normalized to $\gamma_{s,\alpha}$, which is different for each of the curves. This normalization makes the $e^{2 \gamma_{s,\alpha,k} t}$ line the same for all $\alpha$. 
$||e^{\mathbf{B}_{k,\alpha} t}||^2$ displays large non-exponential growth at early times that increases with decreasing $\alpha$.}
\label{max_en_growth}
\end{figure}

@@

\beqar
\label{subs_fourier_n}
n = \sum_{k} n_k \ {\rm sin} \left( k_x x \right) e^{i k_y y + i k_z z} \\
\label{subs_fourier_phi}
\phi = \sum_{k} \phi_k \ {\rm sin} \left( k_x x \right) e^{i k_y y + i k_z z},
\eeqar
where, again, $k$ stands for $(k_x,k_y,k_z)$.
Then, we multiply Eq.~\ref{subs_fourier_n} by $n^*_{k'} \ {\rm sin} \left( k'_x x \right) e^{-i k'_y y - i k'_z z}$ and Eq.~\ref{subs_fourier_phi} by 
$-\phi^*_{k'} \ {\rm sin} \left( k'_x x \right) e^{-i k'_y y - i k'_z z}$, take the volume integral and add the two equations together. The result is

\beqar
\label{Explicit_En_eqn}
\frac{1}{2} \diff{}{t} \left( |n_k|^2 + k^2_\perp |\phi_k|^2 \right) = -i k_y \kappa \phi_k n^*_k - \xi k_z^2 |n_k - \phi_k|^2 \nonumber \\
- D k^4_\perp \left( |n_k|^2 + k^2_\perp |\phi_k|^2 \right) + \sum_{k'} T(k,k'). \quad \quad
\eeqar
Recall that the energy $E_k$ is $\frac{1}{2} \left( |n_k|^2 + k^2_\perp |\phi_k|^2 \right)$, so the left hand side is the time derivative of the energy.
The term $\sum_{k'} T(k,k')$ represents the nonlinear contribution (three-wave transfers between different $k$'s), 
which we do not write explicitly. This energy evolution equation may be written symbolically as

\beq
\label{dEdt_def}
\diff{E_k}{t} = \diff{E_{l,k}}{t} + \diff{E_{nl,k}}{t}
\eeq
where $\diff{E_{l,k}}{t}$ represents the the linear terms on the RHS of Eq.~\ref{Explicit_En_eqn}, which are quadratic in the fluctuating quantities $n_k$ and $\phi_k$.
$\diff{E_{l,k}}{t}$ describes the injection of energy into the fluctuations from the free energy in the equilibrium gradients plus the dissipation from collisions and hyperdiffusion/viscosity.
$\diff{E_{nl,k}}{t}$ represents $\sum_{k'} T(k,k')$ in  Eq.~\ref{Explicit_En_eqn} and comes from the nonlinear terms in Eqs.~\ref{n_eq} and~\ref{phi_eq}. 
The terms in $T(k,k')$ are triadic in the fluctuating quantities and account for the energy exchange between fluctuations with different $k$'s. 
They are also conservative when summed over all wavenumbers: $\sum_{k} \diff{E_{nl,k}}{t} = \sum_{k,k'} T(k,k') = 0$~\cite{camargo1995}.

@@

\beqar
\label{steady_state}
\gamma_k = \gamma_{l,k} + \gamma_{nl,k} = 0, \\
\gamma_k \equiv  \displaystyle\lim_{T \to \infty} \frac{1}{T} \int_0^T \frac{dE_k/dt}{2 E_k} dt \nonumber \\
= \displaystyle\lim_{T \to \infty} \frac{{\rm{Log}}\left[ E_k(T)/E_k(0) \right]}{2 T} = 0.
\eeqar
$\gamma_k$ vanishes, as indicated, for steady or quasi-steady state turbulence, meaning $\gamma_{l,k} = - \gamma_{nl,k}$.

>>>>

\title{Non-Modal Predictions of Turbulent Properties Applied to the Hasegawa-Wakatani Model}

@@

Non-modal analysis generally employs tools such as pseudospectra and maximum growth curves, which have been used to successfully explain
why turbulence is subcritical in various instances~\cite{trefethen1993,camargo1998}. 
However, they have not been used to predict onset criteria in subcritical systems (e.g. the transition Reynold's number in pipe flows).
This paper outlines and illustrates a non-modal method to make quantitative predictions of turbulent properties including turbulent onset criteria. 
Our approach is to define a non-modal growth rate that is analogous to an eigenmode growth rate, but which takes into account transient growth effects.

@@

In the analysis and numerical simulations, which we perform with the BOUT++ code~\cite{dudson2009}, we use periodic boundary conditions in
the $y$ and $z$ directions with zero-value boundaries in the $x$ direction. We use a domain size of $64 \rho_s \times 64 \rho_s \times 320 \rho_s$.
For the 2D version of the model, we replace $-\xi \gradpar^2$ with a single parameter $\alpha$ (defined as $\alpha = 4 \pi k_z^2 \rho_s^2 \omega_{ce} /\nu_{ei})$, 
which is commonly known as the adiabaticity parameter~\cite{camargo1995,camargo1998}. Additionally we use a domain size of $50 \rho_s \times 50 \rho_s$.
In numerical simulations, the turbulence can be suppressed due to the build-up of $k_y=0$ density and potential structures~\cite{biskamp1995}. To prevent this, we subtract out the $k_y=0$
density and potential components, which is the equivalent of adding source terms to Eqs.~\ref{n_eq} and~\ref{phi_eq}~\cite{friedman2012b}.

@@

\beqar
\label{fourier_eqn}
\frac{\partial}{\partial t} \left( \begin{array}{cc} n_k \\ \phi_k \end{array} \right) = \mathbf{A}_k \left( \begin{array}{cc} n_k \\ \phi_k \end{array} \right) + \mathbf{N}_k \left( \begin{array}{cc} n_k \\ \phi_k \end{array} \right), \\ \nonumber \\
\label{A_k}
\mathbf{A}_k = \left( \begin{array}{cc} -\xi k_z^2 - D k_\perp^4 & -i \kappa k_y + \xi k_z^2 \\  \xi \frac{k_z^2}{k_\perp^2} & - \xi \frac{k_z^2}{k_\perp^2} - D k_\perp^4\end{array} \right)
\eeqar
with $\mathbf{A}_k$ being the linear coupling matrix and $\mathbf{N}_k$ the nonlinear operator, which we don't write explicitly.
Now in assessing the degree of normality of the HW model, we must make a choice of norms because one can always choose a norm that makes a particular matrix normal. However, it is much more revealing to choose
a physically motivated norm, especially the energy-norm, than one that simply makes a given matrix normal~\cite{camargo1998,schmid2007,camporeale2010}. Since the energy in a given Fourier mode is

@@

\beqar
\label{fourier_en_eqn}
\frac{\partial}{\partial t} \left( \begin{array}{cc} n_k \\ k_\perp \phi_k \end{array} \right) = \mathbf{B}_k \left( \begin{array}{cc} n_k \\ k_\perp \phi_k \end{array} \right) + \mathbf{N}_k \left( \begin{array}{cc} n_k \\ k_\perp \phi_k \end{array} \right), \\ \nonumber \\
\label{B_k}
\mathbf{B}_k = \left( \begin{array}{cc} -\xi k_z^2 - D k_\perp^4 & -i \kappa \frac{k_y}{k_\perp} + \xi \frac{k_z^2}{k_\perp} \\  \xi \frac{k_z^2}{k_\perp} & - \xi \frac{k_z^2}{k_\perp^2} - D k_\perp^4\end{array} \right)
\eeqar
The square $L_2$-norm of the state vector $\frac{1}{\sqrt{2}} \left( n_k , k_\perp \phi_k \right)^T$ gives the energy contained in the mode $k$.
Therefore, we choose to cast the problem in terms of Eq.~\ref{fourier_en_eqn} rather than Eq.~\ref{fourier_eqn} and focus on the linear matrix $\mathbf{B}_k$ rather than $\mathbf{A}_k$. Thus, we can use
the well-known $L_2$-norm rather than a problem-dependent energy norm but still recover the same results as if we used an energy-norm with the formulation of Eq.~\ref{fourier_eqn}.

In the adiabatic limit (as long as $k_\perp$ remains finite):

@@

\beq
\label{lin_HW}
\diff{u_k}{t} = \mathbf{B}_k u_k
\eeq
where $u_k = \frac{1}{\sqrt{2}} \left( n_k, k_\perp \phi_k \right)^T$.
The solution to this is $u_k(t) = e^{\mathbf{B}_k t} u_k(0)$ where the exponential of the matrix is defined in terms of
its Taylor expansion. The solution is dependent on the initial condition $u_k(0)$.
The energy, which is synonymous with the energy amplification factor since we set $||u_k(0)||^2 =1$ evolves as

\beq
\label{Et}
E_k(t) = ||u_k(t)||^2 = ||e^{\mathbf{B}_k t} u_k(0)||^2
\eeq
Different initial conditions provide optimal energy growth at different times, but the upper envelope of all such energy evolution curves is given by
$E_{k,\rm{max}}(t) = ||e^{\mathbf{B}_k t}||^2$, which does not depend on an initial condition.
Moreover, $||e^{\mathbf{B}_k t}||^2 \ge e^{2 \gamma_{s,k} t}$, where $\gamma_{s,k}$ is the real part of
the fastest growing (or least damped) eigenmode of $\mathbf{B}_k$ (and of $\mathbf{A}_k$, since $\mathbf{B}_k$ is just a similarity transform of $\mathbf{A}_k$~\cite{camargo1998}). 
The equality holds when $\mathbf{B}_k$ is normal. 
Thus, non-normality in practice allows a system to undergo more energy growth than is dictated by the most unstable linear eigenmode. We show this in Fig.~\ref{max_en_growth} where we plot
$||e^{\mathbf{B}_k t}||^2$ for different values of $\alpha$ along with $e^{2 \gamma_{s,k} t}$, showing that transient non-modal growth is strong for small values of the scalar normality measure $1-H$, which scales monotonically
with the adiabaticity parameter $\alpha$.

\begin{figure}
\centerline{\includegraphics[width=0.45\textwidth]{max_en_growth}}
\caption{$E_{k,\alpha,\rm{max}}(t) = ||e^{\mathbf{B}_{k,\alpha} t}||^2$ for $k_x=0, k_y=1, \kappa=1$ for different values of the 2D adiabaticity parameter 
$\alpha$ compared to the energy growth $e^{2 \gamma_{s,\alpha,k} t}$ of the least stable eigenmode of  $\mathbf{B}_{k,\alpha}$. Also indicated for each matrix is its scalar normality measure $1-H$.
The time axis is normalized to $\gamma_{s,\alpha}$, which is different for each of the curves. This normalization makes the $e^{2 \gamma_{s,\alpha,k} t}$ line the same for all $\alpha$. 
$||e^{\mathbf{B}_{k,\alpha} t}||^2$ displays large non-exponential growth at early times that increases with decreasing $\alpha$.}
\label{max_en_growth}
\end{figure}

@@

\beqar
\label{subs_fourier_n}
n = \sum_{k} n_k \ {\rm sin} \left( k_x x \right) e^{i k_y y + i k_z z} \\
\label{subs_fourier_phi}
\phi = \sum_{k} \phi_k \ {\rm sin} \left( k_x x \right) e^{i k_y y + i k_z z},
\eeqar
where, again, $k$ stands for $(k_x,k_y,k_z)$.
Then, we multiply Eq.~\ref{n_eq} by $n^*_{k'} \ {\rm sin} \left( k'_x x \right) e^{-i k'_y y - i k'_z z}$ and Eq.~\ref{phi_eq} by 
$-\phi^*_{k'} \ {\rm sin} \left( k'_x x \right) e^{-i k'_y y - i k'_z z}$, take the volume integral and add the two equations together. The result is

\beqar
\label{Explicit_En_eqn}
\frac{1}{2} \diff{}{t} \left( |n_k|^2 + k^2_\perp |\phi_k|^2 \right) = -i k_y \kappa \phi_k n^*_k - \xi k_z^2 |n_k - \phi_k|^2 \nonumber \\
- D k^4_\perp \left( |n_k|^2 + k^2_\perp |\phi_k|^2 \right) + \sum_{k'} T(k,k'). \quad \quad
\eeqar
Recall that the energy $E_k$ is $\frac{1}{2} \left( |n_k|^2 + k^2_\perp |\phi_k|^2 \right)$, so the left hand side is the time derivative of the energy.
The term $\sum_{k'} T(k,k')$ represents the nonlinear contribution (three-wave transfers between different $k$'s), 
which we do not write explicitly. This energy evolution equation may be written formally as

\beqar
\label{dEdt_def}
\diff{E_k}{t} = \diff{E_{l,k}}{t} + \diff{E_{nl,k}}{t}, \quad {\rm where}\\
\label{dEl_dt_def}
\diff{E_{l,k}}{t} = -i k_y \kappa \phi_k n^*_k - \xi k_z^2 |n_k - \phi_k|^2 \nonumber \\
- D k^4_\perp \left( |n_k|^2 + k^2_\perp |\phi_k|^2 \right) \\
\label{dEnl_dt_def}
 \diff{E_{nl,k}}{t} = \sum_{k'} T(k,k')
\eeqar
$\diff{E_{l,k}}{t}$ is composed of the linear terms on the RHS of Eq.~\ref{Explicit_En_eqn}, which are quadratic in the fluctuating quantities $n_k$ and $\phi_k$.
It describes the injection of energy into the fluctuations from the free energy in the equilibrium gradients plus the dissipation from collisions and hyperdiffusion/viscosity.
$\diff{E_{nl,k}}{t}$ represents $\sum_{k'} T(k,k')$ in  Eq.~\ref{Explicit_En_eqn} and comes from the nonlinear terms in Eqs.~\ref{n_eq} and~\ref{phi_eq}. 
The terms in $T(k,k')$ are triadic in the fluctuating quantities and account for the energy exchange between fluctuations with different $k$'s. 
They are also conservative when summed over all wavenumbers: $\sum_{k} \diff{E_{nl,k}}{t} = \sum_{k,k'} T(k,k') = 0$~\cite{camargo1995}.

@@

\beqar
\label{steady_state}
\gamma_{l,k} + \gamma_{nl,k} = \displaystyle\lim_{T \to \infty} \frac{1}{T} \int_0^T \frac{dE_k/dt}{2 E_k} dt = 0, \\
\label{gamma_l_def}
{\rm where}, \ \gamma_{l,k} \equiv \displaystyle\lim_{T \to \infty} \frac{1}{T} \int_0^T \frac{dE_{l,k}/dt}{2 E_k} dt \\
\label{gamma_nl_def}
 \gamma_{nl,k} \equiv \displaystyle\lim_{T \to \infty} \frac{1}{T} \int_0^T \frac{dE_{nl,k}/dt}{2 E_k} dt 
\eeqar
Thus, assuming steady or quasi-steady state turbulence $\gamma_{l,k} = - \gamma_{nl,k}$. Note that these, and all other growth rates, are defined to be time-independent (time-averaged) quantities.
At the scales of energy injection, $\gamma_{l,k}$ is positive and can be considered the turbulence analog of the linear eigenmode growth rate $\gamma_{s,k}$~\cite{friedman2012b,terry2006b}. 
In fact, $\gamma_{l,k} = \gamma_{s,k}$ if $n_k/\phi_k$ calculated from the nonlinear simulation is constant in time and equal to $n_k/\phi_k$ calculated from the fastest growing eigenmode at $k$. 
In general, though, these ratios are not the same because the turbulence at $k$ is composed of some linear combination of the \emph{two} eigenmodes.
